---
title: "4_1_figure"
author: "Mei"
date: "2024-11-02"
output: pdf_document
---

## Set up work space
```{r}
# Load packages
library(dplyr)
library(tidyverse)
library(broom)
library(ggeffects)
library(ggplot2)

```

## Import data
```{r, include=F}
library(dplyr)
# For Office PC
setwd("C:/Users/kuo.355/OneDrive - The Ohio State University/1_Second_Year_Paper/3_Data_Analysis/Data")
# For Macbook
#setwd("/Users/kuomico/Library/CloudStorage/OneDrive-TheOhioStateUniversity/1_Second_Year_Paper/3_Data_Analysis/Data")

# Save orig data
dta_orig <- readRDS("Cleaned_Data/R/df_sum_long.rds")

# Save used data
dta <- dta_orig

```

## Sample & Variable Selection 
```{r}
df <- dta
nrow(df)
summary(df)


df <- df %>%
  # Keep only the last observation of each id
  group_by(id) %>%
  slice_max(order_by = age_BA, n = 1, with_ties = FALSE) %>% 
  ungroup() %>% 
  # Keep ids with valid college major
  filter(major_apst != "Nonoe or not yet declared") %>%
  mutate(major_apst = droplevels(major_apst)) %>%
  # gen gender X major
  mutate(gendermajor = as.factor(paste(major_apst,gender))) %>%
  # Select used variables
  select(id, gender, race, major_apst, gendermajor, age_BA, year, cum_shift_pos, cum_shift_neg) 
  
summary(df)

# Save data
dta <- df

```

## Distribution, mean and se of Vars
# Fx ~ Fx
In particular, all descriptive statistics should be graphically displayed (50 points). Entire distributions of outcomes and key independent variables are expected. Summaries (means and sds) of controls can be graphed as well.
```{r, eval=F, include=F}
# DV: shift_pos, shift_neg
# IV: gender, race, major_apst, age_BA

#########################
# DV 
########################

# Var: cum_shift_pos, cum_shift_neg
list_x <- c("cum_shift_pos", "cum_shift_neg")

# Histogram
df <- dta
for (x in list_x){
  
  p<-ggplot(data=df,mapping=aes_string(x=x)) + 
  geom_histogram(bins=10,position="dodge") +
  theme_classic()+
  scale_x_continuous(breaks = seq(min(0), max(30), by = 1))

  show(p)
  
}


#######################
# IV
#######################

# Var: gender, race, major_apst
list_x <- c("gender", "major_apst", "race" )

# Plotting
df <- dta

for (x in list_x){
  
  p <- ggplot(data=df, aes_string(x=x, fill=x)) +
  geom_bar(position="dodge",mapping=aes(y=..count..)) +
  theme_classic() +
  scale_fill_brewer(palette = "Dark2") +
  labs(x=x,y="count") +
  theme(legend.position="bottom")
  
show(p)
  
}

# For plotting gender X college_major
df <- dta

ggplot(df, aes(x = major_apst, fill = gender)) + 
  geom_bar(position = "dodge") +  
  theme_classic() +
  labs(x = "Major", y = "Count", fill = "Gender") +
  scale_fill_brewer(palette = "Dark2") 


# Var: age_BA
list_x <- c("age_BA")

# Histogram
df <- dta
for (x in list_x){
  
  p<-ggplot(data=df,mapping=aes_string(x=x)) + 
  geom_histogram(bins=10,position="dodge") +
  theme_classic()+
  scale_x_continuous(breaks = seq(min(0), max(30), by = 1))

  show(p)
  
}


```

## Bivariate relationship between Xs and Ys.
# Fx ~ Fx
```{r}
# IV: cum_shift_pos, cum_shift_neg
# Key IV: gender, major_apst

# Initialize list_plot
list_plot <- list()
df <- dta

# Sample: All but missing in edu_degree
list_var <- c("cum_shift_pos", "cum_shift_neg")
list_var_lab <- c("# Positive EMP Transitions", "# Negative EMP Transitions")

for (i in 1:length(list_var)) {
  
  var <- list_var[[i]]
  var_lab <- list_var_lab[[i]]
  
  p <- ggplot(data=df,
              aes(x = major_apst, y = .data[[var]], fill = major_apst)) + 
    geom_bar(stat = "summary", fun = "mean") + 
    geom_errorbar(stat = "summary", fun.data = "mean_se", fun.args = list(mult = 1.96)) +
    stat_summary(
      aes(label = round(..y.., 3)), 
      geom = "text", 
      fun = mean, 
      vjust = 1.5, 
      color = "white", 
      size = 3.5
    ) +
    facet_wrap(~gender) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = var_lab)  + scale_fill_brewer(palette = "Dark2")
  
  show(p)
  list_plot[[var]] <- p
  
}


```

## Coefficient plots & Margins plot
# Fx ~ Fx
Students must present the results from some version of a linear model (e.g., OLS, logit, poisson, etc.) as a coefficient plot (50 points).
Students should also illustrate the effect of key independent variable(s) using plots of margins from the regression models (40 points).
```{r}
require(ggeffects)

#######################
# Positive 
#######################

# model building
df <- dta
m <- lm(cum_shift_pos ~  age_BA + race + gender + major_apst + gender*major_apst, data=df)
summary(m)

# coef plot
tab_m <- tidy(m, conf.int=TRUE) # Transform the model into a table
tab_m  <- tab_m [-1,] # Remove the intercept

# Make the coefficient plot
p <- ggplot(tab_m ,mapping = aes(x=term,y=estimate,ymin=conf.low,ymax=conf.high)) +
     geom_pointrange() + 
     theme_classic() + coord_flip() + 
     labs(x="Independent Variable",y="# Positive EMP Transitions_OLS Regression Coefficient") + 
     geom_hline(yintercept=0)

show(p)

# marginsplot
p2 <- plot(ggemmeans(m,c("major_apst", "gender"))) + 
  scale_colour_manual("gender", values = c("darkgreen", "darkorange")) +
p2

#######################
# Negative 
#######################

# model building
df <- dta
m <- lm(cum_shift_neg ~  age_BA + race + gender + major_apst + gender*major_apst, data=df)
summary(m)

# coef plot
tab_m <- tidy(m, conf.int=TRUE) # Transform the model into a table
tab_m  <- tab_m [-1,] # Remove the intercept

# Make the coefficient plot
p <- ggplot(tab_m ,mapping = aes(x=term,y=estimate,ymin=conf.low,ymax=conf.high)) +
     geom_pointrange() + 
     theme_classic() + coord_flip() + 
     labs(x="Independent Variable",y="# Negative EMP Transitions_OLS Regression Coefficient") + 
     geom_hline(yintercept=0)

show(p)

# marginsplot
p2 <- plot(ggemmeans(m,c("major_apst", "gender"))) + 
  scale_colour_manual("gender", values = c("darkgreen", "darkorange")) +
p2

```


