---
title: "3_1_desc_stats"
author: "Mei"
date: "2024-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## library
library(knitr)
library(tidyverse)
library(haven)
library(dplyr)
library(ggplot2)
library(glm2)
library(lme4)
library(lmerTest)
library(readr)
library(lubridate)

```

## Import data
```{r, include = F}
# Library
library(haven)
#library(lavaan)
library(tidyverse)

# Import from dta. 
# This will lead to unmatched class of vars with ggplot2.
# setwd("C:/Users/kuo.355/OneDrive - The Ohio State University/1_Second_Year_Paper/3_Data_Analysis/Data")
# dta_orig <- read_dta("Cleaned_Data/step1245_wide.dta")
# dta <- dta_orig

## Try import as csv.
## Get rid of labels and relabel from here.

# For Office PC
setwd("C:/Users/kuo.355/OneDrive - The Ohio State University/1_Second_Year_Paper/3_Data_Analysis/Data")
dta_orig_week <- read.csv("Cleaned_Data/step5_wide.csv")
dta_orig_year <- read.csv("Cleaned_Data/step12345_wide.csv")

# For Macbook
# setwd("/Users/kuomico/Library/CloudStorage/OneDrive-TheOhioStateUniversity/1_Second_Year_Paper/3_Data_Analysis/Data")
# dta_orig <- read.csv("Cleaned_Data/step12345_wide.csv")

# list for storing df_merge
list_merge <- list()

```

##### PART. 0, Set up parameters for Slicing data ##### 
## Re-Run
```{r set cutpoint parameters}

# set the cut point
cutpoint <- c(seq(0, 8000, by = 1000), nrow(dta_orig_week))

# each subset is from (cutpoint[n]+1) to cutpoint[n+1]
# num of reps
ns <- 1:(length(cutpoint) - 1)
ns
n <- 9 # Should be 1~9
cutpoint[n]+1
```

##### PART. 1, Generating Weekly Long-Form Data #####

## Reshape from wide to long ## Re-Run
```{r, include=F}
## Start from id == 1
# Initialize
wide.data <- dta_orig_week

# Sort the column by column names and no.of week
wide.data <- wide.data %>% select(order(names(wide.data)))

# Renames rows and columns
list_var <- c("id", "job_id", "job_wage", "job_allwage", "job_hour", "job_full", "job_status",
              "job_occ", "job_ind", "job_8egp_cat", "job_12egp_cat")


# Each pass of the loop, it creates a nested data file (long.data2) and then appends that 
# to the bottom of long form data.

# Reshape the data by cut-point, and store them to separate dataframe in the list
# observation period: 1997/w1(1924)~2022/w40(3263)
list_long <- list()

  for(i in (cutpoint[n]+1):cutpoint[n+1]){
    print(i)

    df <- data.frame(

     # Time-invariant variables: gender, race, p_edu, edu_degree, citizen, fam_inc, asvab, and degree dates
    id = wide.data[i, which(names(wide.data) == "id")],
     week_job_id = t(wide.data[i, which(names(wide.data) == "week_job_id_1924"):which(names(wide.data) == "week_job_id_3263")]),
    week_job_wage = t(wide.data[i, which(names(wide.data) == "week_job_wage_1924"):which(names(wide.data) =="week_job_wage_3263")]),
    week_job_allwage = t(wide.data[i, which(names(wide.data) == "week_job_allwage_1924"):which(names(wide.data) == "week_job_allwage_3263")]),
    week_job_hour = t(wide.data[i, which(names(wide.data) == "week_job_hour_1924"):which(names(wide.data) == "week_job_hour_3263")]),
    week_job_full = t(wide.data[i, which(names(wide.data) == "week_job_full_1924"):which(names(wide.data) == "week_job_full_3263")]),
     week_job_status = t(wide.data[i, which(names(wide.data) == "week_job_status_1924"):which(names(wide.data) == "week_job_status_3263")]),

    week_job_occ = t(wide.data[i, which(names(wide.data) == "week_job_occ_1924"):which(names(wide.data) == "week_job_occ_3263")]),
    week_job_ind = t(wide.data[i, which(names(wide.data) == "week_job_ind_1924"):which(names(wide.data) == "week_job_ind_3263")]),
    week_job_8egp = t(wide.data[i, which(names(wide.data) == "week_job_8egp_1924"):which(names(wide.data) == "week_job_8egp_3263")]),
    week_job_12egp = t(wide.data[i, which(names(wide.data) == "week_job_12egp_1924"):which(names(wide.data) == "week_job_12egp_3263")])

  )
    rownames(df) <- NULL
    colnames(df)<- list_var
    
    
    list_long[[i]] <- df
    if (i > (cutpoint[n]+1) & i <= cutpoint[n+1]){
      list_long[[i]] <- rbind(list_long[[i-1]],list_long[[i]])
    }
    
  }
  


# Store long-form data
long.data <- list_long[[cutpoint[n+1]]]

# Check num of ids and observations.
range(unique(long.data$id))
length(unique(long.data$id)) # should = 8984 (ids)
nrow(long.data)/length(unique(long.data$id)) # should = 1340 (cweeks)

# Assign cw
nid <- length(unique(as.factor(long.data$id)))
w1 <- 1924
wn <- 3263
long.data$cw <- rep(w1:wn, times = nid)

# Save data
dta <- long.data
#saveRDS(dta, "Cleaned_Data/R/long_emp.rds")

# Clear the data
rm(list_long)

```

## Rename and re-label vars ## Re-Run
```{r, include=F}
# Initialize
df <- long.data

# Key variables 
colnames(df)
 # c("id", "gender",  "race", "p_edu", "edu_degree", "citizen", 
 #              "fam_inc", "asvab", "bdate", "date_fm",
 #              "date_HS", "date_AA", "date_BA", "date_MA", "date_PHD", "date_PROF", 
 #              "major_field", "major_app","major_stem", "major_apst",
 #              "child_res", "child_nonres", "child_all", "census_msa", "census_region", "enroll", "weight",
 #              "job_wage", "job_allwage", "job_hour", "job_full", 
 #              "job_occ", "job_ind", "job_8egp_cat", "job_12egp_cat", "job_id", "job_sat_cat", "cw")

# 2: Replace all blanks("") with NA; replace missing with NAs.
df[1,"job_ind"]
typeof(df[1,"job_ind"])
df[df == ""] <- NA
df[df$job_id %in% c(5, -4, 4), "job_id"] <- NA
  
# 3: Numerize cont. vars & date vars
list_var <- c("job_wage", "job_allwage", "job_hour")

# Print the type of each variable in the list
var_types <- sapply(list_var, function(var) class(df[[var]]))
print(var_types)

# adjust wage <- wage/100
summary(df$job_wage)
df$job_wage <- df$job_wage/100
summary(df$job_allwage)
df$job_allwage <- df$job_allwage/100

# 4: Factorize cat. vars.
list_var <- c("job_full", "job_ind", "job_occ", "job_8egp_cat", "job_12egp_cat","job_id","job_status")

# Replace missing with NAs
summary(df$job_full)
df$job_full[df$job_full == "Missing"] <- NA 
df$job_full[df$job_full == "Not employed"] <- "Not employed"

# Replace invalid job_id & job_hour
df$job_id[df$job_id < 9000 ] <- NA 
df$job_hour[df$job_hour < 0 ] <- NA 

# Print the type of each variable in the list
var_types <- sapply(list_var, function(var) class(df[[var]]))
print(var_types)

# Factorize
df <- df %>% mutate(across(list_var, as.factor))
typeof(df$job_full)
summary(df$job_full)
typeof(df$id)

# 5: Gen diff. types of  EGP & job_sat variables.
df <- df %>% mutate(
  job_8egp = recode(job_8egp_cat,
                    "Class 1" = 1,
                    "Class 2" = 2,
                    "Class 3" = 3,
                    "Class 4" = 4,
                    "Class 5" = 5,
                    "Class 6" = 6,
                    "Class 7" = 7,
                    "Military" = 8
                    ),
  job_12egp = recode(job_12egp_cat,
                     "Class 1" = 1,
                    "Class 2" = 2,
                    "Class 3a" = 3,
                    "Class 3b" = 4,
                    "Class 4a" = 5,
                    "Class 4b" = 6,
                    "Class 4c" = 7,
                    "Class 5" = 8,
                    "Class 6" = 9,
                    "Class 7a" = 10,
                    "Class 7b" = 11,
                    "Military" = 12),
  job_full_bi = recode(job_full,
                   "Full-time" = "Employed", "Part-time" = "Employed",
                    "Not employed" = "Not employed"
                   )
)

df$job_full_bi <- as.factor(df$job_full_bi)
table(df$job_full, df$job_full_bi)
table(df$job_8egp_cat, df$job_8egp)

# 6: Level ordinal cat. vars.
# List: "p_edu", "edu_degree", "job_full", "job_12egp", "job_sat"
df <- df %>% mutate(
  job_full_lv = factor(job_full, levels=c("Not employed","Part-time","Full-time"), order = T),
  job_12egp_lv = factor(job_12egp, levels=c(1,2,3,4,5,6,7,8,9,10,11,12), order = T),
  job_8egp_lv = factor(job_8egp, levels=c(1,2,3,4,5,6,7,8), order = T)
)

summary(df$job_8egp_lv)
levels(df$job_8egp_lv)

# generate year according to cw
list_year <- 1997:2022
list_year_first_week <- c(1924, 1976, 2028, 2080, 2132, 2184, 2236, 2288, 2340, 2392, 2444, 2496, 2548, 2600,
                          2652, 2704, 2756, 2808, 2860, 2912, 2964, 3016, 3068, 3120, 3172, 3224)
df <- df %>%
  mutate(year = list_year[findInterval(cw, list_year_first_week)])

# Check
summary(df)

# Save data
long.data_cleaned <- df

```

##### PART. 2, Gen Yearly Long-Form Data by Identifying Main Job #####

## Aggregate weekly emp history to yearly history. ## Re-Run 
```{r reshape and aggregate data}
# Observation period: 1997/w1 ~ 2022/w40
# Var to be generated: job_id, job_hour, job_full, job_status, job_wage, job_allwage, job_8egp, job_12egp, job_ind

# Initialize
df <- long.data_cleaned
list_sum <- list()

########################
# Identify the Main job
########################

# Step 2: Sum the total work hours of each job_id in each year. Order it.
df_sum <- df %>%
  group_by(id, year, job_id) %>%
  summarize(job_year_hour = sum(job_hour, na.rm = T), 
            count = n()) %>%
  arrange(desc(job_year_hour)) %>%
  mutate(job_order = row_number())

# Step 3: Keep the main job and merge it with occupation, wage, and industry.
df_sum <- df_sum %>% 
  filter(job_order == 1) %>%
  arrange(id, year)

# Step 4: Save the data
list_sum[["job_id"]] <- df_sum

##############################
# Compute Yearly Working Hours
##############################

# Step 2: Sum the total work hours (including every job) of each year. 
df_sum <- df %>%
  group_by(id, year) %>%
  summarize(
    job_year_hour = sum(job_hour, na.rm = TRUE),
    job_n_unemp_outofLF = sum(job_status %in% c( "1","2", "3", "4", "5", "6"), na.rm = TRUE),
    job_n_skip_missing = sum(job_status %in% c("0", "-1", "-2", "-4", "-5"), na.rm = TRUE),
    count = n()
  ) %>%
  mutate(
    job_full = case_when( 
      job_year_hour/count >= 35 ~ "Full-time",
      job_year_hour/count > 0 & job_year_hour/count < 35 ~ "Part-time",
      job_year_hour == 0 ~ "Not employed",
      is.na(job_year_hour) ~ "Not employed"
    ),
    job_avg_uemp_outofLF = round(job_n_unemp_outofLF/count,2), 
    job_avg_skip_missing = round(job_n_skip_missing/count,2)
    ) 

# Step 3: Factorize and check the dist. of job_full
df_sum$job_full <- as.factor(df_sum$job_full)
summary(df_sum)

# Step 4: Save the data
list_sum[["job_full"]] <- df_sum

```

## Gen Yearly Interview & Demographic Data
```{r int_date_year}
# Initialize
dta <- dta_orig_year 

# Sort the data
dta <- dta %>%
  select(order(names(dta))) %>%
  arrange(id)

# reshape 
list_var <- c("id", "gender",  "race", "p_edu", "edu_degree", "citizen", 
              "fam_inc", "asvab", "bdate", "date_fm",
              "date_HS", "date_AA", "date_BA", "date_MA", "date_PHD", "date_PROF", 
              "major_field", "major_app","major_stem", "major_apst",
              "fm_status", "child_res", "child_nonres", "child_all", "census_msa", "census_region", "enroll", "weight", "intdate")

wide.data <- dta
for (i in 1:nrow(wide.data)){
  print(i)
  df_long_append <- data.frame(
    id = wide.data[i, which(names(wide.data) == "id")],
    gender = wide.data[i, which(names(wide.data) == "gender")],
    race = wide.data[i, which(names(wide.data) == "race")],
    p_edu = wide.data[i, which(names(wide.data) == "p_edu")],
    edu_degree = wide.data[i, which(names(wide.data) == "edu_degree")],
    citizen = wide.data[i, which(names(wide.data) == "citizen")],
    fam_inc = wide.data[i, which(names(wide.data) == "fam_inc")],
    asvab = wide.data[i, which(names(wide.data) == "asvab")],
    
    bdate = wide.data[i, which(names(wide.data) == "bdate")],
    date_fm = wide.data[i, which(names(wide.data) == "fm_date")],
    degree_date_HS = wide.data[i, which(names(wide.data) == "degree_date_HS")],
    degree_date_AA = wide.data[i, which(names(wide.data) == "degree_date_AA")],
    degree_date_BA = wide.data[i, which(names(wide.data) == "degree_date_BA")],
    degree_date_MA = wide.data[i, which(names(wide.data) == "degree_date_MA")],
    degree_date_PHD = wide.data[i, which(names(wide.data) == "degree_date_PHD")],
    degree_date_PROF = wide.data[i, which(names(wide.data) == "degree_date_PROF")],
    
    college_major_field = wide.data[i, which(names(wide.data) == "college_major_field_")],
    college_major_app = wide.data[i, which(names(wide.data) == "college_major_app_")],
    college_major_stem = wide.data[i, which(names(wide.data) == "college_major_stem_")],
    college_major_apst = wide.data[i, which(names(wide.data) == "college_major_apst")],
    
    # Dynamic ranges for the time-varying variables
    fam_status = t(wide.data[i, which(names(wide.data) == "fm_status_1997"):which(names(wide.data) == "fm_status_2022")]),
    child_res = t(wide.data[i, which(names(wide.data) == "child_res_1997"):which(names(wide.data) == "child_res_2022")]),
    child_nonres = t(wide.data[i, which(names(wide.data) == "child_nonres_1997"):which(names(wide.data) == "child_nonres_2022")]),
    child_all = t(wide.data[i, which(names(wide.data) == "child_all_1997"):which(names(wide.data) == "child_all_2022")]),
    census_msa = t(wide.data[i, which(names(wide.data) == "census_msa_1997"):which(names(wide.data) == "census_msa_2022")]),
    census_region = t(wide.data[i, which(names(wide.data) == "census_region_1997"):which(names(wide.data) == "census_region_2022")]),
    enroll_college = t(wide.data[i, which(names(wide.data) == "enroll_1997"):which(names(wide.data) == "enroll_2022")]),
    weight = t(wide.data[i, which(names(wide.data) == "weight_1997"):which(names(wide.data) == "weight_2022")]), 
    intdate = t(dta[i, which(names(dta) == "intdate_1997"):which(names(dta) == "intdate_2022")])
  )
  rownames(df_long_append) <- NULL
  colnames(df_long_append)<- list_var
  
  if(i == 1){
    df <- df_long_append
    
  }else{
    df <- rbind(df, df_long_append)
  }
  
}

colnames(df)<- list_var
    
# check num of ids and years
length(unique(df$id)) # n = 8984
nrow(df)/length(unique(df$id)) # 26 years

# paste the year
years <- rep(1997:2022, times = length(unique(df$id)))
df$year <- years

# save data
df_control_year <- df

```

## Save orig year control data
```{r}
# Initialize
df <- df_control_year

# Save yearly long-form data
write.csv(df, "Cleaned_Data/R/df_control_year.csv", row.names = FALSE)
saveRDS(df, "Cleaned_Data/R/df_control_year.rds")

```

## Import data, if needed
```{r}
df_control_year <- readRDS("Cleaned_Data/R/df_control_year.rds")

```

## Clean type and names of vars
```{r Clean type and names of vars}
# Initialize
df <- df_control_year
# Fix date and time-varying age 
 df <- df %>% 
    mutate(bdate = year(as.Date(bdate, origin = "1960-01-01")),
           age_bdate = round(year - bdate),
           age_BA= round(year - date_BA))
  summary(df %>% select(id, bdate, date_BA, year, age_bdate, age_BA))  


  
# Fill the blanks with NA
df[df == ""] <- NA
df[df$enroll < 0, "enroll"] <- NA

# Factorize cat. variable
list_var <- c("gender", "race", "citizen", "p_edu", "edu_degree",
              "census_region", "census_msa", "enroll", "fm_status",
              "major_field", "major_app", "major_stem", "major_apst")
  # Print the type of each variable in the list
  var_types <- sapply(list_var, function(var) class(df[[var]]))
  print(var_types)
  
  # Fix census_msa
  df <- df %>%
  mutate(census_msa = case_when(
    census_msa == 1 ~ "MSA",
    census_msa == 0 ~ "Non-MSA",
    TRUE ~ NA_character_
  ))

  
  # Factorize.
  df$fm_status[df$fm_status == ""] <- "Missing"
  df$fm_status[is.na(df$fm_status)] <- "Missing"
  df <- df %>% mutate(across(list_var, as.factor))
  typeof(df$major_app)
  summary(df$major_apst)
  typeof(df$id)  
  

# Final check
summary(df)
# check num of ids and years
length(unique(df$id)) # n = 8984
nrow(df)/length(unique(df$id)) # 26 years

# Save data
df_control_year <- df

```


## Construct time-varying vars 
```{r, include=F}
# Initialize
df <- df_control_year

# Gen time-varying status based on date_vars.
list_var <- c("date_fm", "date_HS", "date_AA", "date_BA", 
              "date_MA", "date_PHD", "date_PROF")

df <- df %>%
  mutate(across(all_of(list_var), 
                ~ ifelse(year >= .x & !is.na(.x), 1, ifelse(year <.x & !is.na(.x), 0, NA)), 
                .names = "flag_{col}"))

df <- df %>%
  mutate(
    # Create the time-varying highest degree variable
    flag_highest = case_when(
      edu_degree %in% c("< HS", "Missing") ~ 1,  # "< HS" or "Missing" get assigned 1
      
      # Apply the logic for those with BA+ (BA or higher)
      edu_degree == "BA+" & flag_date_BA == 1 ~ 1,
      edu_degree == "BA+" & flag_date_BA == 0 ~ 0,  # Not attained BA+ yet
      edu_degree == "BA+" & is.na(flag_date_BA) & flag_date_HS == 1 ~ 1,
      edu_degree == "BA+" & is.na(flag_date_BA) & flag_date_HS == 0 ~ 0,
      edu_degree == "BA+" & is.na(flag_date_BA) & is.na(flag_date_HS) ~ 1, # For missing in degree_date_HS & BA
      
      # For BA graduates
      edu_degree == "BA" & flag_date_BA == 1 ~ 1,
      edu_degree == "BA" & flag_date_BA == 0 ~ 0,  # Not attained BA yet
      edu_degree == "BA" & is.na(flag_date_BA) & flag_date_HS == 1 ~ 1,
      edu_degree == "BA" & is.na(flag_date_BA) & flag_date_HS == 0 ~ 0,
      edu_degree == "BA" & is.na(flag_date_BA) & is.na(flag_date_HS) ~ 1, 
      
      # For AA graduates
      edu_degree == "AA" & flag_date_AA == 1 ~ 1,
      edu_degree == "AA" & flag_date_AA == 0 ~ 0,  # Not attained AA yet
      edu_degree == "AA" & is.na(flag_date_AA) & flag_date_HS == 1 ~ 1,
      edu_degree == "AA" & is.na(flag_date_AA) & flag_date_HS == 0 ~ 0,
      edu_degree == "AA" & is.na(flag_date_AA) & is.na(flag_date_HS) ~ 1, 
      
      # For HS graduates
      edu_degree == "HS/GED" & flag_date_HS == 1 ~ 1,
      edu_degree == "HS/GED" & flag_date_HS == 0 ~ 0,  # Not attained HS yet
      edu_degree == "HS/GED" & is.na(flag_date_HS) ~ 1,

      TRUE ~ 0  # Default case
    )
  )

# Generate age
df <- df %>% 
  mutate(age = year - bdate)

# Drop the date vars
# list_var <-  c("bdate", "date_fm", "date_HS", "date_AA", "date_BA", 
#               "date_MA", "date_PHD", "date_PROF")
# 
# df <- df[, !names(df) %in% list_var]

# Check
summary(df)

# Check time-varying BA age
summary(df %>%
            filter(edu_degree %in% c("BA", "BA+"), 
            major_apst %in% c("Academic non-STEM", "Academic STEM",
                              "Applied non-STEM", "Applied STEM"),
            flag_highest == 1) %>%
            
            select(id, year, age, age_BA) ## age_BA should >= 0, as selecting on date after BA.
        )  

# Save data
df_control_year <- df

```
Note
1. Those with 
a. missing date in both their highest degree & high school graduation
b. highest edu level is either < HS or missing,
are observed across their available measurements.

## Merge summary data with original data ## Re-Run
```{r merge}
# Initialize
df_week <- long.data_cleaned
df_job_id <- list_sum[["job_id"]]
df_job_full <- list_sum[["job_full"]]
df_control_year <- df_control_year

# Keep used vars
df_week <- df_week %>% 
  select(id, year, job_id, job_wage, job_allwage, job_occ,
         job_8egp, job_8egp_lv, job_12egp, job_12egp_lv, job_ind) %>%
  group_by(id, year, job_id) %>%
  slice(1) %>%
  ungroup()

df_job_id <- df_job_id %>%
  select(id, year, job_id)

# Merge two yearly sum data, and then link them with emp info in weekly raw data.
df <- merge(df_job_id, df_job_full, by = c("id", "year"))
df <- merge(df, df_week, by = c("id", "year", "job_id"))

# Merge with yearly control data
df <- merge(df, df_control_year, by = c("id", "year"))

# Check num of ids and observations.
range(unique(df$id))
length(unique(df$id)) # should = 1000 (ids)
nrow(df)/length(unique(df$id)) # should = 26 (years)

# Save data
df_merge <- df
# store the subset df_merge into list
list_merge[[n]] <- df_merge

```


## Combine df_merge ## Final
```{r combine summary_and_merge_df}
# rbind all the stored long-form data
indices <- 1:9
df <- do.call(rbind, list_merge[indices])

# check num of ids and years
length(unique(df$id)) # n = 8984
nrow(df)/length(unique(df$id)) # 26 years

# save data
df_combine_long <- df
```

## Export data
```{r, eval=F, include=F}
# Initialize
df <- df_combine_long

# Save yearly long-form data
write.csv(df, "Cleaned_Data/R/long_year.csv", row.names = FALSE)
saveRDS(df, "Cleaned_Data/R/long_year.rds")

```




















