---
title: "3_1_desc_stats"
author: "Mei"
date: "2024-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## library
library(knitr)
library(tidyverse)
library(haven)
library(dplyr)
library(ggplot2)
library(glm2)
library(lme4)
library(lmerTest)
library(readr)
library(lubridate)

```

## Import data
```{r, include = F}
# Library
library(haven)
#library(lavaan)
library(tidyverse)

# Import from dta. 
# This will lead to unmatched class of vars with ggplot2.
# setwd("C:/Users/kuo.355/OneDrive - The Ohio State University/1_Second_Year_Paper/3_Data_Analysis/Data")
# dta_orig <- read_dta("Cleaned_Data/step1245_wide.dta")
# dta <- dta_orig

## Try import as csv.
## Get rid of labels and relabel from here.

# For Office PC
setwd("C:/Users/kuo.355/OneDrive - The Ohio State University/1_Second_Year_Paper/3_Data_Analysis/Data")
dta_orig <- read.csv("Cleaned_Data/step12345_wide.csv")
# For Macbook
# setwd("/Users/kuomico/Library/CloudStorage/OneDrive-TheOhioStateUniversity/1_Second_Year_Paper/3_Data_Analysis/Data")
# dta_orig <- read.csv("Cleaned_Data/step12345_wide.csv")


```

## Reshape from wide to long
```{r, include=F}
## Start from id == 1
# Initialize
wide.data <- dta_orig

# Sort the column by column names
wide.data <- wide.data %>% select(order(names(wide.data)))

# Reshape the data
long.data <- data.frame(
  
  # Time-invariant variables: gender, race, p_edu, edu_degree, citizen, fam_inc, asvab, and degree dates
  id = wide.data[1, which(names(wide.data) == "id")],
  gender = wide.data[1, which(names(wide.data) == "gender")],
  race = wide.data[1, which(names(wide.data) == "race")],
  p_edu = wide.data[1, which(names(wide.data) == "p_edu")],
  edu_degree = wide.data[1, which(names(wide.data) == "edu_degree")],
  citizen = wide.data[1, which(names(wide.data) == "citizen")],
  fam_inc = wide.data[1, which(names(wide.data) == "fam_inc")],
  asvab = wide.data[1, which(names(wide.data) == "asvab")],
  
  bdate = wide.data[1, which(names(wide.data) == "bdate")],
  date_fm = wide.data[1, which(names(wide.data) == "fm_date")],
  degree_date_HS = wide.data[1, which(names(wide.data) == "degree_date_HS")],
  degree_date_AA = wide.data[1, which(names(wide.data) == "degree_date_AA")],
  degree_date_BA = wide.data[1, which(names(wide.data) == "degree_date_BA")],
  degree_date_MA = wide.data[1, which(names(wide.data) == "degree_date_MA")],
  degree_date_PHD = wide.data[1, which(names(wide.data) == "degree_date_PHD")],
  degree_date_PROF = wide.data[1, which(names(wide.data) == "degree_date_PROF")],
  
  college_major_field = wide.data[1, which(names(wide.data) == "college_major_field_")],
  college_major_app = wide.data[1, which(names(wide.data) == "college_major_app_")],
  college_major_stem = wide.data[1, which(names(wide.data) == "college_major_stem_")],
  college_major_apst = wide.data[1, which(names(wide.data) == "college_major_apst")],
  
  # Dynamic ranges for the time-varying variables
  child_res = t(wide.data[1, which(names(wide.data) == "child_res_1924"):which(names(wide.data) == "child_res_3263")]),
  child_nonres = t(wide.data[1, which(names(wide.data) == "child_nonres_1924"):which(names(wide.data) == "child_nonres_3263")]),
  child_all = t(wide.data[1, which(names(wide.data) == "child_all_1924"):which(names(wide.data) == "child_all_3263")]),
  census_msa = t(wide.data[1, which(names(wide.data) == "census_msa_1924"):which(names(wide.data) == "census_msa_3263")]),
  census_region = t(wide.data[1, which(names(wide.data) == "census_region_1924"):which(names(wide.data) == "census_region_3263")]),
  enroll_college = t(wide.data[1, which(names(wide.data) == "enroll_college_1924"):which(names(wide.data) == "enroll_college_3263")]),
  weight = t(wide.data[1, which(names(wide.data) == "weight_1924"):which(names(wide.data) == "weight_3263")]), 
  
  week_job_wage = t(wide.data[1, which(names(wide.data) == "week_job_wage_1924"):which(names(wide.data) == "week_job_wage_3263")]),
  week_job_allwage = t(wide.data[1, which(names(wide.data) == "week_job_allwage_1924"):which(names(wide.data) == "week_job_allwage_3263")]),
  week_job_hour = t(wide.data[1, which(names(wide.data) == "week_job_hour_1924"):which(names(wide.data) == "week_job_hour_3263")]),
  week_job_full = t(wide.data[1, which(names(wide.data) == "week_job_full_1924"):which(names(wide.data) == "week_job_full_3263")]),
  
  week_job_occ = t(wide.data[1, which(names(wide.data) == "week_job_occ_1924"):which(names(wide.data) == "week_job_occ_3263")]),
  week_job_ind = t(wide.data[1, which(names(wide.data) == "week_job_ind_1924"):which(names(wide.data) == "week_job_ind_3263")]),
  week_job_8egp = t(wide.data[1, which(names(wide.data) == "week_job_8egp_1924"):which(names(wide.data) == "week_job_8egp_3263")]),
  week_job_12egp = t(wide.data[1, which(names(wide.data) == "week_job_12egp_1924"):which(names(wide.data) == "week_job_12egp_3263")]),
  week_job_id = t(wide.data[1, which(names(wide.data) == "week_job_id_1924"):which(names(wide.data) == "week_job_id_3263")]),
  week_job_sat = t(wide.data[1, which(names(wide.data) == "week_job_sat_1924"):which(names(wide.data) == "week_job_sat_3263")])
  
)

# Renames rows and columns
list_var <- c("id", "gender",  "race", "p_edu", "edu_degree", "citizen", 
              "fam_inc", "asvab", "bdate", "date_fm",
              "date_HS", "date_AA", "date_BA", "date_MA", "date_PHD", "date_PROF", 
              "major_field", "major_app","major_stem", "major_apst",
              "child_res", "child_nonres", "child_all", "census_msa", "census_region", "enroll", "weight",
              "job_wage", "job_allwage", "job_hour", "job_full", 
              "job_occ", "job_ind", "job_8egp_cat", "job_12egp_cat", "job_id", "job_sat_cat")

rownames(long.data) <- NULL
colnames(long.data) <- list_var

# Each pass of the loop, it creates a nested data file (long.data2) and then appends that 
# to the bottom of long form data.
for(i in 2:100){
  print(i)
  long.data2 <- data.frame(
  
  # Time-invariant variables: gender, race, p_edu, edu_degree, citizen, fam_inc, asvab, and degree dates
  id = wide.data[i, which(names(wide.data) == "id")],
  gender = wide.data[i, which(names(wide.data) == "gender")],
  race = wide.data[i, which(names(wide.data) == "race")],
  p_edu = wide.data[i, which(names(wide.data) == "p_edu")],
  edu_degree = wide.data[i, which(names(wide.data) == "edu_degree")],
  citizen = wide.data[i, which(names(wide.data) == "citizen")],
  fam_inc = wide.data[i, which(names(wide.data) == "fam_inc")],
  asvab = wide.data[i, which(names(wide.data) == "asvab")],
  
  bdate = wide.data[i, which(names(wide.data) == "bdate")],
  date_fm = wide.data[i, which(names(wide.data) == "fm_date")],
  degree_date_HS = wide.data[i, which(names(wide.data) == "degree_date_HS")],
  degree_date_AA = wide.data[i, which(names(wide.data) == "degree_date_AA")],
  degree_date_BA = wide.data[i, which(names(wide.data) == "degree_date_BA")],
  degree_date_MA = wide.data[i, which(names(wide.data) == "degree_date_MA")],
  degree_date_PHD = wide.data[i, which(names(wide.data) == "degree_date_PHD")],
  degree_date_PROF = wide.data[i, which(names(wide.data) == "degree_date_PROF")],
  
  college_major_field = wide.data[i, which(names(wide.data) == "college_major_field_")],
  college_major_app = wide.data[i, which(names(wide.data) == "college_major_app_")],
  college_major_stem = wide.data[i, which(names(wide.data) == "college_major_stem_")],
  college_major_apst = wide.data[i, which(names(wide.data) == "college_major_apst")],
  
  # Dynamic ranges for the time-varying variables
  child_res = t(wide.data[i, which(names(wide.data) == "child_res_1924"):which(names(wide.data) == "child_res_3263")]),
  child_nonres = t(wide.data[i, which(names(wide.data) == "child_nonres_1924"):which(names(wide.data) == "child_nonres_3263")]),
  child_all = t(wide.data[i, which(names(wide.data) == "child_all_1924"):which(names(wide.data) == "child_all_3263")]),
  census_msa = t(wide.data[i, which(names(wide.data) == "census_msa_1924"):which(names(wide.data) == "census_msa_3263")]),
  census_region = t(wide.data[i, which(names(wide.data) == "census_region_1924"):which(names(wide.data) == "census_region_3263")]),
  enroll_college = t(wide.data[i, which(names(wide.data) == "enroll_college_1924"):which(names(wide.data) == "enroll_college_3263")]),
  weight = t(wide.data[i, which(names(wide.data) == "weight_1924"):which(names(wide.data) == "weight_3263")]),
  week_job_wage = t(wide.data[i, which(names(wide.data) == "week_job_wage_1924"):which(names(wide.data) == "week_job_wage_3263")]),
  week_job_allwage = t(wide.data[i, which(names(wide.data) == "week_job_allwage_1924"):which(names(wide.data) == "week_job_allwage_3263")]),
  week_job_hour = t(wide.data[i, which(names(wide.data) == "week_job_hour_1924"):which(names(wide.data) == "week_job_hour_3263")]),
  week_job_full = t(wide.data[i, which(names(wide.data) == "week_job_full_1924"):which(names(wide.data) == "week_job_full_3263")]),
  
  week_job_occ = t(wide.data[i, which(names(wide.data) == "week_job_occ_1924"):which(names(wide.data) == "week_job_occ_3263")]),
  week_job_ind = t(wide.data[i, which(names(wide.data) == "week_job_ind_1924"):which(names(wide.data) == "week_job_ind_3263")]),
  week_job_8egp = t(wide.data[i, which(names(wide.data) == "week_job_8egp_1924"):which(names(wide.data) == "week_job_8egp_3263")]),
  week_job_12egp = t(wide.data[i, which(names(wide.data) == "week_job_12egp_1924"):which(names(wide.data) == "week_job_12egp_3263")]),
  week_job_id = t(wide.data[i, which(names(wide.data) == "week_job_id_1924"):which(names(wide.data) == "week_job_id_3263")]),
  week_job_sat = t(wide.data[i, which(names(wide.data) == "week_job_sat_1924"):which(names(wide.data) == "week_job_sat_3263")])
  
)
  rownames(long.data2) <- NULL
  colnames(long.data2)<- list_var
  long.data <- rbind(long.data,long.data2)
}

# Assign cw 
nid <- length(unique(as.factor(long.data$id)))
w1 <- 1924
wn <- 3263
long.data$cw <- rep(w1:wn, times = nid)  

# Save data
dta <- long.data
saveRDS(dta, "Cleaned_Data/R/long_full.rds")

```

## Rename and re-label vars
```{r, include=F}
# Initialize
df <- dta

# Key variables 
colnames(df)
 # c("id", "gender",  "race", "p_edu", "edu_degree", "citizen", 
 #              "fam_inc", "asvab", "bdate", "date_fm",
 #              "date_HS", "date_AA", "date_BA", "date_MA", "date_PHD", "date_PROF", 
 #              "major_field", "major_app","major_stem", "major_apst",
 #              "child_res", "child_nonres", "child_all", "census_msa", "census_region", "enroll", "weight",
 #              "job_wage", "job_allwage", "job_hour", "job_full", 
 #              "job_occ", "job_ind", "job_8egp_cat", "job_12egp_cat", "job_id", "job_sat_cat", "cw")

# 2: Replace all blanks("") with NA; replace missing with NAs.
df[1,"job_occ"]
typeof(df[1,"job_occ"])
df[df == ""] <- NA

df[df$job_id %in% c(5, -4, 4), "job_id"] <- NA
df <- df %>%
  mutate(major_apst = ifelse(!(edu_degree %in% c("BA", "BA+")), NA, major_apst))
table(df$edu_degree, df$major_apst)
  
# 3: Numerize cont. vars & date vars
list_var <- c("fam_inc", "asvab", "bdate", "date_fm",
               "date_HS", "date_AA", "date_BA", "date_MA", "date_PHD", "date_PROF",
               "child_res", "child_nonres", "child_all", "weight", 
               "job_wage", "job_allwage", "job_hour")

# Print the type of each variable in the list
var_types <- sapply(list_var, function(var) class(df[[var]]))
print(var_types)

# adjust wage <- wage/100
summary(df$job_wage)
df$job_wage <- df$job_wage/100
summary(df$job_allwage)
df$job_allwage <- df$job_allwage/100

# 4: Factorize cat. vars.
list_var <- c("gender", "race", "citizen", "p_edu", "edu_degree",
              "census_region", "census_msa", "enroll",
              "major_field", "major_app", "major_stem", "major_apst",
              "job_full", "job_occ", "job_ind", "job_8egp_cat", "job_12egp_cat","job_id","job_sat_cat")

# Replace missing with NAs
summary(df$job_full)
df$job_full[df$job_full == "Missing"] <- NA 
df$job_full[df$job_full == "Not employed"] <- "Not-employed"

# Print the type of each variable in the list
var_types <- sapply(list_var, function(var) class(df[[var]]))
print(var_types)

# Factorize
df <- df %>% mutate(across(list_var, as.factor))
typeof(df$major_app)
summary(df$major_app)
typeof(df$id)

# 5: Gen diff. types of  EGP & job_sat variables.
df <- df %>% mutate(
  job_8egp = recode(job_8egp_cat,
                    "Class 1" = 1,
                    "Class 2" = 2,
                    "Class 3" = 3,
                    "Class 4" = 4,
                    "Class 5" = 5,
                    "Class 6" = 6,
                    "Class 7" = 7,
                    "Military" = 8
                    ),
  job_12egp = recode(job_12egp_cat,
                     "Class 1" = 1,
                    "Class 2" = 2,
                    "Class 3a" = 3,
                    "Class 3b" = 4,
                    "Class 4a" = 5,
                    "Class 4b" = 6,
                    "Class 4c" = 7,
                    "Class 5" = 8,
                    "Class 6" = 9,
                    "Class 7a" = 10,
                    "Class 7b" = 11,
                    "Military" = 12),
  job_sat = recode(job_sat_cat,
                   "Dislike it very much" = 1, "Dislike it somewhat" = 2, "Think it is OK" = 3,
                   "Like it fairly well" = 4, "Like it very much" = 5
                   ),
  job_full_bi = recode(job_full,
                   "Full-time" = "Employed", "Part-time" = "Employed",
                   "Missing" = NA, "Not employed" = "Not-employed"
                   )
)

df$job_full_bi <- as.factor(df$job_full_bi)
table(df$job_8egp_cat, df$job_8egp)
table(df$job_sat_cat, df$job_sat)

# 6: Level ordinal cat. vars.
# List: "p_edu", "edu_degree", "job_full", "job_12egp", "job_sat"
df <- df %>% mutate(
  p_edu_lv = factor(p_edu, levels=c("Missing","<HS","HS","Some College","BA or more"), order = T),
  edu_degree_lv = factor(edu_degree, levels=c("Missing","< HS","HS/GED","AA","BA","BA+"), order = T),
  job_full_lv = factor(job_full, levels=c("Not employed","Part-time","Full-time"), order = T),
  job_12egp_lv = factor(job_12egp, levels=c(1,2,3,4,5,6,7,8,9,10,11,12), order = T),
  job_8egp_lv = factor(job_8egp, levels=c(1,2,3,4,5,6,7,8), order = T),
  job_sat_lv = factor(job_sat, levels=c(1,2,3,4,5), order = T)
  
)

summary(df$p_edu_lv)
levels(df$p_edu_lv)

# Check
summary(df)

# Save data
dta <- df

```

## Construct time-invariant vars
```{r, include=F}
# Initialize
df <- dta

# gender, race, edu interaction terms
df <- df %>% mutate(genderrace = as.factor(paste(race, gender)),
                    gendermajor = as.factor(paste(major_apst, gender)),
                    racemajor = as.factor(paste(major_apst, race)),
                    genderedu = as.factor(paste(edu_degree, gender)),
                    raceedu = as.factor(paste(edu_degree, race))
                    )

# gen 

summary(df)
dta <- df

```

## Construct time-varying vars 
Note
1. Those with 
a. missing date in both their highest degree & high school graduation
b. highest edu level is either < HS or missing,
are observed across their available measurements.
```{r, include=F}
# Initialize
df <- dta

# Gen time-varying status based on date_vars.
list_var <- c("date_fm", "date_HS", "date_AA", "date_BA", 
              "date_MA", "date_PHD", "date_PROF")

df <- df %>%
  mutate(across(all_of(list_var), 
                ~ ifelse(cw >= .x & !is.na(.x), 1, ifelse(cw <.x & !is.na(.x), 0, NA)), 
                .names = "flag_{col}"))

df <- df %>%
  mutate(
    # Create the time-varying highest degree variable
    flag_highest = case_when(
      edu_degree %in% c("< HS", "Missing") ~ 1,  # "< HS" or "Missing" get assigned 1
      
      # Apply the logic for those with BA+ (BA or higher)
      edu_degree == "BA+" & flag_date_BA == 1 ~ 1,
      edu_degree == "BA+" & flag_date_BA == 0 ~ 0,  # Not attained BA+ yet
      edu_degree == "BA+" & is.na(flag_date_BA) & flag_date_HS == 1 ~ 1,
      edu_degree == "BA+" & is.na(flag_date_BA) & flag_date_HS == 0 ~ 0,
      edu_degree == "BA+" & is.na(flag_date_BA) & is.na(flag_date_HS) ~ 1, # For missing in degree_date_HS & BA
      
      # For BA graduates
      edu_degree == "BA" & flag_date_BA == 1 ~ 1,
      edu_degree == "BA" & flag_date_BA == 0 ~ 0,  # Not attained BA yet
      edu_degree == "BA" & is.na(flag_date_BA) & flag_date_HS == 1 ~ 1,
      edu_degree == "BA" & is.na(flag_date_BA) & flag_date_HS == 0 ~ 0,
      edu_degree == "BA" & is.na(flag_date_BA) & is.na(flag_date_HS) ~ 1, 
      
      # For AA graduates
      edu_degree == "AA" & flag_date_AA == 1 ~ 1,
      edu_degree == "AA" & flag_date_AA == 0 ~ 0,  # Not attained AA yet
      edu_degree == "AA" & is.na(flag_date_AA) & flag_date_HS == 1 ~ 1,
      edu_degree == "AA" & is.na(flag_date_AA) & flag_date_HS == 0 ~ 0,
      edu_degree == "AA" & is.na(flag_date_AA) & is.na(flag_date_HS) ~ 1, 
      
      # For HS graduates
      edu_degree == "HS/GED" & flag_date_HS == 1 ~ 1,
      edu_degree == "HS/GED" & flag_date_HS == 0 ~ 0,  # Not attained HS yet
      edu_degree == "HS/GED" & is.na(flag_date_HS) ~ 1,

      TRUE ~ 0  # Default case
    )
  )

# Drop the date vars
df <- df[, !names(df) %in% list_var]

# Check
summary(df)

# Save data
dta <- df

```

## Sample selection
```{r}
# df <- dta
# Original sample size: id = 8,984, id-week = ###### (~8984*52*28 = 13080704).
# nrow(df)
# length(unique(df$id))

# Schultz: drop id that have <15y of work history.

```

## Export data
```{r, eval=F, include=F}
# Save long-form data
df <- dta
write.csv(df, "Cleaned_Data/R/long.csv", row.names = FALSE)
saveRDS(df, "Cleaned_Data/R/long.rds")

# Save wide-form data
df <- filter(dta, cw == 1924) # keep the first obs of each id
write.csv(df, "Cleaned_Data/R/wide.csv", row.names = FALSE)
saveRDS(df, "Cleaned_Data/R/wide.rds")

```

