---
title: "3_1_desc_stats"
author: "Mei"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## library
library(knitr)
library(tidyverse)
library(haven)
library(dplyr)
library(ggplot2)
library(glm2)
library(lme4)
library(lmerTest)
library(readr)
library(lubridate)

```

## Import data
```{r Import data}
# For office-PC
setwd("C:/Users/kuo.355/OneDrive - The Ohio State University/1_Second_Year_Paper/3_Data_Analysis/Data")

# For Macbook
#setwd("/Users/kuomico/Library/CloudStorage/OneDrive-TheOhioStateUniversity/1_Second_Year_Paper/3_Data_Analysis/Data")
# Load data
dta_orig_long <- readRDS("Cleaned_Data/R/long_year.rds")
df_control_year <- readRDS("Cleaned_Data/R/df_control_year.rds")

```
 
## Create alternative datadest: distinguish non-interviewed from not-employed 
```{r alternatvie dataset}
# Initialize
df <- dta_orig_long

# Indicator of never employed in observation period
df_never <- df %>%
  group_by(id) %>%
  summarise(sum_job_year_hour = sum(job_year_hour, na.rm = TRUE)) %>%
  mutate(job_never = ifelse(sum_job_year_hour == 0, 1, 0)) %>%
  select(id, job_never) # Keep only the relevant columns

  # Merge the indicator back into the original dataframe to preserve the structure
  df <- df %>%
    left_join(df_never, by = "id")

  
  
# check prop. of not interviewed years
na_proportion <- df %>% 
  summarize(proportion_na = mean(is.na(intdate)))
na_proportion 

  ## 17.1% of person-years are not interviewed.

# Distinguish Not-employed people: (job_full == "Not employed")
  # 1) not employed : > 50% in job_avg_uemp_outofLF; > 50% in enroll
  # 2) interviewed but skip/missing : job_avg_skip_missing
  # 3) not interviewed : intdate

na_proportion <- df %>% 
  summarize(proportion_na = sum(job_full == "Not employed")/nrow(df))
na_proportion 

  ## 31.03% person-year are Not employed.
na_proportion <- df %>% 
  filter(job_full == "Not employed") %>%
  summarize(proportion_uemp = mean(job_avg_uemp_outofLF), 
            proportion_skip = mean(job_avg_skip_missing),
            proportion_enroll = sum(enroll == "Enrolled", na.rm = T)/n(),
            proportion_notint = mean(is.na(intdate)))
na_proportion 

  ## Among these not-emp person-years, 53% not employed/out of LF, 42.9% skip, 5% enroll, 36% not interviewed.

#######################
# WARNING: DROP SAMPLE
#######################

# Exclude the person-years of 
# 1) not employed 2) emp/out of LF < 50% 3) not enrolled 4) not interviewed year.
drop_sample <- df %>% 
  filter(job_full == "Not employed" & 
         job_avg_uemp_outofLF < .5 &
         enroll != "Enrolled" &
         is.na(intdate)) 
nrow(drop_sample)/nrow(df)

  ## 9% person-year will be removed.

df <- df %>%
   filter(!(job_full == "Not employed" & 
         job_avg_uemp_outofLF < .5 &
         enroll != "Enrolled" &
         is.na(intdate)))
nrow(df)/nrow(dta_orig_long)

  ## 90.68% person-year remains.

# Gen a count of valid waves across id
df <- df %>%
  group_by(id) %>%
  mutate(valid_wave = n()) %>%
  ungroup

# Check
summary(df)
# Save data
dta <- df

```

## Check sample selection process (Don't run unless necessarily.)
# Part 1 criteria: 1) Valid BA dates, 2) >= 2 valid waves
# Part 2 criteria: 1) Based on part 1, 2) >= 3 valid waves, 3) >=2 part-time/full-time waves with valid wages and EGP-class.
```{r}
df <- dta %>%
  
   # Sampling on 1) BA degrees holders 2) with valid college majors 3) years after attaining BA
  filter(edu_degree %in% c("BA", "BA+"), 
         !is.na(date_BA), 
         major_apst %in% c("Academic non-STEM", "Academic STEM", "Applied non-STEM", "Applied STEM"),
         flag_highest == 1) %>%
  
  # gen a interaction var to check
  mutate(gendermajor = paste(gender, major_apst)) %>%
  
  group_by(id) %>%
  mutate(valid_wave = n()) %>%
  ungroup

# Check #, gen X major of IDs that are never employed.
## These ppl are controlled in part 1 analysis, and are excluded in part 2 analysis.
table(df$job_never, df$gendermajor) ## 6 person-year for men; 8 person-year for women

test <- df %>%
  group_by(id) %>%
  filter(year == min(year)) %>%
  ungroup

table(test$job_never, test$gendermajor) ## 1 man; 1 woman


# Check #, gen X major of IDs that have only 1 valid wave, regardless of their EMP status in this wave.
## These ppl are supposed to be removed from both part 1 & part 2 analysis.
test <- df %>%
  group_by(id) %>%
  filter(year == min(year)) %>%
  mutate(wave1 = ifelse(valid_wave == 1, 1, 0)) %>%
  ungroup 
  
table(test$wave1, test$gendermajor) ## 3 man; 5 woman


```


############# X --> M: Gender X Edu Variation in Career Movements ### ## Cross-sectional ####################

## Compute the overall # of movements
```{r, include=F}
# Locate data
df <- dta 

# Sampling on 1) BA degrees holders 2) with valid college majors 3) years after attaining BA
list_var <- c("edu_degree", "major_apst", "flag_date_BA")
summary(df %>% select(all_of(list_var)))
table(df$edu_degree, df$major_apst) ## 312 + 364 person-years are missing in major

############################################################################################################
# Sample: Long-form. Obs starts after attaining highest degree. Do not apply for <HS & Missing.   
# Missing (either true missing or non-interviewed is not addressed.)

# Variable to be generated
list_var <- c("id", "valid_wave", "shift_pos", "shift_neg", "shift_all", "n_out")

df_sum <- df %>%
  # sort the data
  arrange(id, year) %>%
  
  # Sampling on 1) BA degrees holders 2) with valid college majors 3) years after attaining BA
  filter(edu_degree %in% c("BA", "BA+"), 
         !is.na(date_BA), 
         major_apst %in% c("Academic non-STEM", "Academic STEM", "Applied non-STEM", "Applied STEM"),
         flag_highest == 1) %>%
  
  # sort
  group_by(id) %>%
  
  # Create a new column with the previous year's job_full status
  mutate(
    prev_job_full = lag(job_full)  
  ) %>%

  # Compare the current emp status with the previous wave
  mutate(
    shift_full_full = case_when(prev_job_full == "Full-time" & job_full == "Full-time" ~ 1, TRUE ~0),
    shift_part_full = case_when(prev_job_full == "Part-time" & job_full == "Full-time" ~ 1, TRUE ~0),
    shift_non_full = case_when(prev_job_full == "Not employed" & job_full == "Full-time" ~ 1, TRUE ~0),
    shift_non_part = case_when(prev_job_full == "Not employed" & job_full == "Part-time" ~ 1, TRUE ~0),
    
    shift_full_part = case_when(prev_job_full == "Full-time" & job_full == "Part-time" ~ 1, TRUE ~0),
    shift_full_non = case_when(prev_job_full == "Full-time" & job_full == "Not employed" ~ 1, TRUE ~0),
    shift_part_non = case_when(prev_job_full == "Part-time" & job_full == "Not employed" ~ 1, TRUE ~0),
    shift_non_non = case_when(prev_job_full == "Not employed" & job_full == "Not employed" ~ 1, TRUE ~0),
    
    shift_part_part = case_when(prev_job_full == "Part-time" & job_full == "Part-time" ~ 1, TRUE ~0)
    
  ) %>%
  
  summarise(
    n_shift_full_full = sum(shift_full_full, na.rm = TRUE),
    n_shift_part_full = sum(shift_part_full, na.rm = TRUE),
    n_shift_non_full = sum(shift_non_full, na.rm = TRUE),
    n_shift_non_part = sum(shift_non_part, na.rm = TRUE),
    
    n_shift_full_part = sum(shift_full_part, na.rm = TRUE),
    n_shift_full_non = sum(shift_full_non, na.rm = TRUE),
    n_shift_part_non = sum(shift_part_non, na.rm = TRUE),
    n_shift_non_non = sum(shift_non_non, na.rm = TRUE),
    
    n_shift_part_part = sum(shift_part_part, na.rm = TRUE)
    
  ) %>%
  mutate(
    shift_pos = n_shift_full_full + n_shift_part_full + n_shift_non_full + n_shift_non_part,
    shift_neg = n_shift_full_part + n_shift_full_non + n_shift_part_non + n_shift_non_non,
    shift_all = n_shift_part_full + n_shift_non_full + n_shift_non_part + n_shift_full_part + n_shift_full_non + n_shift_part_non,
    
  ) 

##########################################################################################

# Check
summary(df_sum)

# merge it with id's demo characteristics
list_var <- c("id", "gender", "race", "major_apst")
df_sum_cross <- merge(df_sum, dta %>% 
                        arrange(year) %>% 
                        select(all_of(list_var)) %>% 
                        group_by(id) %>% 
                        slice(1), by="id")

# gen interaction terms
df_sum_cross <- df_sum_cross %>% mutate(genderrace = as.factor(paste(race, gender)),
                    gendermajor = as.factor(paste(major_apst, gender)),
                    racemajor = as.factor(paste(major_apst, race))
                    )
# Check
summary(df_sum_cross)

```

## Desc. stats, cross-sectional
```{r x--> M, desc, cross-sectional}
# Initialize
df <- df_sum_cross

list_x <- c("gender", "major_apst", "gendermajor")
list_y <- c("shift_pos", "shift_neg", "shift_all")

for (i in 1:length(list_y)){
  y <-list_y[i]
  
  for (j in 1:length(list_x)){
  x <- list_x[j]
 
  print(paste(x,y))
  df_sum <- df %>% group_by(!!sym(x)) %>% 
    summarize(y_mean = mean(!!sym(y), na.rm = TRUE), y_med = median(!!sym(y), na.rm = TRUE)) 
  
  print(df_sum)
  }
  
}

```

### X --> M: Gender X Edu Variation in Career Movements ### ## Longitudinal
## Compute time-varying Risks indicators
```{r x--> M, risks, longitudinal}
# Library
library(dplyr)

# Locate data
df <- dta 

############################################################################################################
# Sample: Long-form. Obs starts after attaining highest degree. Do not apply for <HS & Missing.   
# Missing (either true missing or non-interviewed is not addressed.)

# Variable to be generated
list_var <- c("id", "shift_pos", "shift_neg", "shift_all", "n_out")

df_sum <- df %>%
  # sort the data
  arrange(id, year) %>%
  
  # Sampling on 1) BA degrees holders 2) with valid college majors 3) years after attaining BA
  filter(edu_degree %in% c("BA", "BA+"), 
         !is.na(date_BA), 
         major_apst %in% c("Academic non-STEM", "Academic STEM", "Applied non-STEM", "Applied STEM"),
         flag_highest == 1) %>%
  
  # group by id
  group_by(id) %>%
  
  # Create a new column with the previous year's job_full status
  mutate(
    prev_job_full = lag(job_full)  
  ) %>%

  # Compare the current emp status with the previous wave
  mutate(
    shift_full_full = case_when(prev_job_full == "Full-time" & job_full == "Full-time" ~ 1, TRUE ~0),
    shift_part_full = case_when(prev_job_full == "Part-time" & job_full == "Full-time" ~ 1, TRUE ~0),
    shift_non_full = case_when(prev_job_full == "Not employed" & job_full == "Full-time" ~ 1, TRUE ~0),
    shift_non_part = case_when(prev_job_full == "Not employed" & job_full == "Part-time" ~ 1, TRUE ~0),
    
    shift_full_part = case_when(prev_job_full == "Full-time" & job_full == "Part-time" ~ 1, TRUE ~0),
    shift_full_non = case_when(prev_job_full == "Full-time" & job_full == "Not employed" ~ 1, TRUE ~0),
    shift_part_non = case_when(prev_job_full == "Part-time" & job_full == "Not employed" ~ 1, TRUE ~0),
    shift_non_non = case_when(prev_job_full == "Not employed" & job_full == "Not employed" ~ 1, TRUE ~0), 
    
    shift_part_part = case_when(prev_job_full == "Part-time" & job_full == "Part-time" ~ 1, TRUE ~0)
    
  ) %>%
  
  # summarize pos and neg movements, respectively
  mutate(
    shift_pos = shift_full_full + shift_part_full + shift_non_full + shift_non_part,
    shift_neg = shift_full_part + shift_full_non + shift_part_non + shift_non_non,
    shift_all = shift_part_full + shift_non_full + shift_non_part + shift_full_part + shift_full_non + shift_part_non,
  ) %>%
  
  # Cumulative sums within each id over time
  mutate(
    cum_shift_pos = cumsum(shift_pos),
    cum_shift_neg = cumsum(shift_neg),
    cum_shift_all = cumsum(shift_all),
    cum_shift_part = cumsum(shift_part_part)
  ) %>%
  
  ungroup() 

##################################################

# check
list_var <- c("id", "year", "prev_job_full", "shift_pos", "shift_neg", "shift_all", "cum_shift_pos", "cum_shift_neg", "cum_shift_all", "cum_shift_part")
summary(df_sum %>% select(all_of(list_var))) # shift_* should range from 0 ~ 1
summary(df_sum)
length(unique(df_sum$id))
# save data
df_sum_long <- df_sum

```

## Export data
```{r}
library(foreign)

# Save cross-sectional data
df <- df_sum_cross
# write.csv(df, "Cleaned_Data/R/sum_wide.csv", row.names = FALSE)
saveRDS(df, "Cleaned_Data/R/df_sum_cross.rds")
write.dta(df, "Cleaned_Data/R/df_sum_cross.dta")

# Save longitudinal data
df <- df_sum_long
# write.csv(df, "Cleaned_Data/R/sum_wide.csv", row.names = FALSE)
saveRDS(df, "Cleaned_Data/R/df_sum_long.rds")
write.dta(df, "Cleaned_Data/R/df_sum_long.dta")

```



