---
title: ""
author: ""
date: ""
output: 
  pdf_document:
     latex_engine: xelatex
  citation_package: biblatex
geometry: "left = 2.5cm, right = 2cm, top = 2cm, bottom = 2cm"
fontsize: 12pt
header-includes:
  - \usepackage{float}
  - \usepackage{setspace}
  - \usepackage{ragged2e}
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE, eval=F}
#tinytex::install_tinytex()

# Ref https://everyday.codes/tutorials/how-to-use-latex-in-rmarkdown/

# Ref https://bookdown.org/yihui/rmarkdown-cookbook/latex-hardcore.html

# https://ourcodingclub.github.io/tutorials/rmarkdown-dissertation/

# github code
# https://github.com/AnnaChirumbolo/dissertation_template_with_rmd/tree/main

```

<!-- Title page -->

\pagenumbering{gobble}

\begin{center}

\vspace{9cm}

\Large
{\bf The Ohio State University}

\Large
{\bf Department of Sociology}


\vspace{5cm}

\Large
{\bf College Education, Employment Divergences, \\and the Gender Wage Gap}

\vspace{1 cm}

\normalsize

By

\vspace{0.5 cm}

\Large

{\bf Mei-Yu Kuo}

\vspace{2cm}

\Large
M.A. Thesis

\vspace{3 cm}

Committee Members: \\
Vincent J. Roscigno (Chair), \\
Hui Zheng, \\
Rachel Dwyer, \\
Fangqi Wen

\vspace{1.5 cm}
\normalsize
January 2025

\end{center}

<!-- Table of Contents -->

\newpage
\begin{center}
\tableofcontents
\end{center}
<!-- Begin of Content -->

\pagenumbering{arabic}


\newpage

```{r abstract, child='1_abstract.Rmd'}
```

\newpage

```{r intro, child = '2_intro.Rmd'}
```

\newpage

```{r literature, child = '3_lit.Rmd'}
```

\newpage

```{r methods, child = '4_method.Rmd'}
```

\newpage

```{r results, child = '5_result.Rmd'}
```

\newpage

```{r discussion, child = '6_discussion.Rmd'}
```


\newpage

```{r appendix, child = '8_table_figure.Rmd'}
```

\newpage

```{r appendix, child = '9_appendix.Rmd'}
```

\newpage

```{r biblio, child = '7_bibliography.Rmd'}
```
